Analyzing Encrypted Web Traffic with Single- and Multi-Class SVMs

### 2011-12-06 report/paper.tex
paper edits

This problem is naturally phrased as a supervised learning 
problem. Given a set of packet traces, where each trace is labelled 
with the website that a user visited to generate it, we would 
like to determine the website that a user visited to generate a new, 
unlabeled trace. The point of training such a classifier is not 
necessarily to make it easier for a censor to do its job, but rather 
to understand the security guarantees that a user gets from using 
an anonymizing proxy to forward encrypted web traffic, as well as 
the effectiveness of proposed defenses against traffic analysis. 
One such proposed defense is to generate a background request to a 
random website for every encrypted request to a sensitive website~\cite{torfingerprinting}. 
This defense, which I will refer to as the random-visit defense, 
attempts to confuse the classifier by the addition of random 
noise to the traffic generated by a visit to a sensitive website.

@@

Below I summarize the tasks that I completed for this project.
\begin{itemize}
\item \textbf{Data generation and preprocessing.} I generated packet 
traces of visits to HTTPS websites, both with and without random 
background visits. From these traces, I generated a feature vector for 
each visit. This process is described in more detail in Section~\ref{sec:data}. 
\item \textbf{Training binary and multiclass SVMs.} I used the 
generated data to train several types of SVMs from the Python scikits-learn 
library~\cite{sklearn}. These included a binary SVM, as well as one-vs-all
and one-vs-one multiclass SVMs based on a binary SVM. I implemented $k$-fold cross-validation 
to determine the SVM parameters.
\item \textbf{Implementing a multiclass SVM and a single-class SVM for anomaly detection.} 
I built a multiclass SVM using the Python \texttt{cvxopt} library~\cite{cvxopt}.
 To investigate the 
open-world setting, I also built and trained a single-class SVM to perform 
anomaly detection.
\item \textbf{Designing and implementing a custom dimensionality reduction algorithm.} 
Because \texttt{cvxopt} does not scale to very large problems, the feature vectors from the 
packet traces were too large to run on the multiclass and single-class SVMs that I built. 
I designed a feature selection algorithm that uses data from website visits without the 
random-visit defense to select features for the data generated using the random-visit defense.
\end{itemize}

@@

I implemented $k$-fold cross-validation to determine parameters of each SVM	, and also to verify to study how 
SVM performance changes as the amount of training data increases. Figure~\ref{fig:vary-visits} shows how the 
one-vs-one multiclass SVM performance increases with the number of samples for each target site. As we expect,
performance increases somewhat linearly with the amount of training data. Figure~\ref{fig:vary-labels} shows 
how one-vs-all multiclass SVM performance deteriorates as the number of classes increases. Though it is difficult 
to extrapolate from the graph, we see a large fall in performance as the number of classes grows from $2$ to $4$, 
and then a slight leveling off. Hopefully, this means that the classifier's performance scales well with the 
number of classes and that these results are transferrable to a more realistic closed-world setting with dozens or 
hundreds of target sites.

@@

For the open world setting, I implemented the dual form of a single-class SVM with kernels, using the 
\texttt{cvxopt} library. This implementation can be found in \texttt{anomaly\_detection.py}. For each target site, I trained 
a separate anomaly detector to recognize visits to that site as positive events, in hopes that visits to any other site 
would register as negative events, thereby enabling the detection of visits to target sites. Unfortunately, the single-class
SVM proved to be of little utility in this setting. Because of \texttt{cvxopt}'s memory limitations, I could only run the 
anomaly detector on dimensionality-reduced multi-visit data. As in the closed-world setting, I performed $k$-fold cross 
validation to tune the parameters, but I unable to find a set of parameters for either a linear or RBF kernel that resulted 
in a high true-positive/true-negative rate and a low false-positive/false-negative rate. In this application, a low true positive 
rate might be acceptable if the false-positive rate is also low: a censoring entity probably does not want to disrupt a large 
percentage of Internet traffic because it has registered as a false positive on an anomaly detector for a target site, but an 
anomaly detector that allows a censor to discover a small percentage of target site visits without disrupting other traffic 
might be desirable. I therefore tuned the parameters for low false positive rates, and with an RBF kernel with $\beta=0.0000005$, $v=0.01$, and multi-visit data with 30 visits per target site, reduced to $130$ dimensions, the anomaly detector obtained an average true positive 
rate of $15\%$ with an average false positive rate of $4\%$. In future, I would like to experiment with the minimum enclosing ball 
anomaly detector to determine if that model better fits this data.

@@

The feature extraction algorithm is parameterized by $b$, a number of blocks, and $\epsilon$, a window around packet lengths.
The algorithm is given as input a feature vector $x$, a matrix of single-visit vectors $X$, and a desired number of 
features $n$. To produce a dimensionality-reduced version of $x$ with $n$ features, the algorithm does the following:
\begin{enumerate}
\item Divide $x$ into $b$ blocks.
\item For each packet $p$ in $x$:
\begin{enumerate}
\item $Score(p) \leftarrow 0$
\item $i \leftarrow $ the index of the block that contains $p$
\item For each vector $x \in X$:
\begin{enumerate}
\item $bx \leftarrow $ the packets in the $i^{th}$ block of $x$
\item $Score(p) \leftarrow Score(p) + \frac{k}{m}$, where $k$ is the number of packets in $bx$ with the same direction as $p$ and with a length within $\epsilon$ of the length of $p$, and $m$ is the total number of packets in $bx$ with the same direction as $p$.
\end{enumerate}
\end{enumerate}
\item Return $x' \leftarrow $ the vector with a feature for the length and direction of each of the $n$ distinct highest scoring packets
\end{enumerate}

In short, the algorithm scores each packet of each vector according to how similar it is to packets in the corresponding blocks of
single-visit feature vectors. Different features are chosen for different vectors. The feature extraction depends only on the 
single-visit data set, not on other multi-visit data points or on the labels.

@@

This algorithm seems to lead to good performance on multiclass SVMs, by picking out the few packet lengths/directions that 
distinguish target site visits. I used $\epsilon=100$ and $b=15$, and found these parameters to work well enough.
 In future, I would like to understand more rigorously why this algorithm works, what good choices of parameters are,
  and if there are
existing, well-studied alternatives that perform the same type of feature selection.

@@

In this project, I investigated the performance of several single- and multiclass SVMs on packet traces from visits to encrypted websites. 
I found that the recently proposed random-visit defense against traffic analysis does not hold up, since a SVM with 
dimensionality-reduced multi-visit data can achieve up to $70\%$ accuracy against this defense. In the future, I would like to better 
understand the feature extraction algorithm that I designed from a rigorous standpoint, and also study how to improve anomaly detection 
for the open-world setting of the encrypted traffic analysis problem.

>>>>

This problem is naturally phrased as a supervised learning 
problem. We train on a set of packet traces, where each trace is labelled 
with the website that a user visited to generate it. Given a new,
unlabeled trace, we would 
like to determine the website that a user visited to generate this trace.
 The point of training such a classifier is not 
necessarily to make it easier for a censor to do its job, but rather 
to understand the security guarantees that a user gets from using 
an anonymizing proxy to forward encrypted web traffic, as well as 
the effectiveness of proposed defenses against traffic analysis. 
One such proposed defense is to generate a background request to a 
random website for every encrypted request to a sensitive website~\cite{torfingerprinting}. 
This defense, which I will refer to as the random-visit defense, 
attempts to confuse the classifier by the addition of random 
noise to the traffic generated by a visit to a sensitive website.

@@

Below I summarize the tasks that I completed for this project.
\begin{itemize}
\item \textbf{Data generation and preprocessing.} I generated packet 
traces of visits to HTTPS websites, both with and without random 
background visits. From these traces, I generated a feature vector for 
each visit. This process is described in more detail in Section~\ref{sec:data}. 
\item \textbf{Training binary and multiclass SVMs.} I used the 
generated data to train several types of SVMs from the Python scikits-learn 
library~\cite{sklearn}. These included a binary SVM, as well as one-vs-all
and one-vs-one multiclass SVMs based on a binary SVM. I implemented $k$-fold cross-validation 
to determine the SVM parameters.
\item \textbf{Implementing a multiclass SVM and a single-class SVM for anomaly detection.} 
I built a multiclass SVM using the Python \texttt{cvxopt} library~\cite{cvxopt}.
 To investigate the 
open-world setting, I also built and trained a single-class SVM to perform 
anomaly detection.
\item \textbf{Designing and implementing a custom dimensionality reduction algorithm.} 
Because \texttt{cvxopt} does not scale to very large problems~\cite{cvxoptscale}, the feature vectors from the 
packet traces were too large to run on the multiclass and single-class SVMs that I built. 
I designed a feature selection algorithm (Section~\ref{sec:dimred}) similar to mutual information 
that uses data from website visits without the 
random-visit defense to select features for the data generated using the random-visit defense.
\end{itemize}

@@

I implemented $k$-fold cross-validation to determine parameters of each SVM, and also to study how 
SVM performance changes as the amount of training data increases. Figure~\ref{fig:vary-visits} shows how the 
one-vs-one multiclass SVM performance increases with the number of samples for each target site. As we expect,
performance increases somewhat linearly with the amount of training data. Figure~\ref{fig:vary-labels} shows 
how one-vs-all multiclass SVM performance deteriorates as the number of classes increases. Though it is difficult 
to extrapolate from the graph, we see a large fall in performance as the number of classes grows from $2$ to $4$, 
and then a slight leveling off. Hopefully, this means that the classifier's performance scales well with the 
number of classes and that these results are transferrable to a more realistic closed-world setting with dozens or 
hundreds of target sites.

@@

For the open world setting, I implemented the dual form of a single-class SVM with kernels, using the 
\texttt{cvxopt} library. This implementation can be found in \texttt{anomaly\_detection.py}. For each target site, I trained 
a separate anomaly detector to recognize visits to that site as positive events, in hopes that visits to any other site 
would register as negative events, thereby enabling the detection of visits to target sites. Unfortunately, the single-class
SVM proved to be of little utility in this setting. Because of \texttt{cvxopt}'s memory limitations, I could only run the 
anomaly detector on dimensionality-reduced multi-visit data. As in the closed-world setting, I performed $k$-fold cross 
validation to tune the parameters, but I was unable to find a set of parameters for either a linear or RBF kernel that resulted 
in a high true-positive/true-negative rate and a low false-positive/false-negative rate. In this application, a low true-positive 
rate might be acceptable if the false-positive rate is also low: a censoring entity probably does not want to disrupt a large 
percentage of Internet traffic because it has registered as a false positive on an anomaly detector for a target site, but an 
anomaly detector that allows a censor to discover a small percentage of target site visits without disrupting other traffic 
might be desirable. I therefore tuned the parameters for low false positive rates, and with an RBF kernel with $\beta=0.0000005$, $v=0.01$, and multi-visit data with 30 visits per target site, reduced to $130$ dimensions, the anomaly detector obtained an average true positive 
rate of $15\%$ with an average false positive rate of $4\%$. In future, I would like to experiment with the minimum enclosing ball 
anomaly detector to determine if that model better fits this data.

@@

The feature extraction algorithm is parameterized by $b$, a number of blocks, and $\epsilon$, a window around packet lengths.
The algorithm is given as input a multi-visit feature vector $x$, a matrix of single-visit vectors $X$, and a desired number of 
features $n$. To produce a dimensionality-reduced version of $x$ with $n$ features, the algorithm does the following:
\begin{enumerate}
\item Divide $x$ into $b$ blocks.
\item For each packet $p$ in $x$:
\begin{enumerate}
\item $Score(p) \leftarrow 0$
\item $i \leftarrow $ the index of the block that contains $p$
\item For each vector $x' \in X$:
\begin{enumerate}
\item Divide $x'$ into $b$ blocks.
\item $bx' \leftarrow $ the packets in the $i^{th}$ block of $x'$
\item $Score(p) \leftarrow Score(p) + \frac{k}{m}$, where $k$ is the number of packets in $bx'$ with the same direction as $p$ and with a length within $\epsilon$ of the length of $p$, and $m$ is the total number of packets in $bx'$ with the same direction as $p$.
\end{enumerate}
\end{enumerate}
\item Return $x_{new} \leftarrow $ the vector with a feature for the length and direction of each of the $n$ distinct highest scoring packets
\end{enumerate}

In short, the algorithm scores each packet of each vector according to how similar it is to packets in the corresponding blocks of
single-visit feature vectors. Different features are chosen for different vectors. The feature extraction depends only on the 
single-visit data set, not on other multi-visit data points or on the labels. This algorithm is akin to mutual information, except 
that it operates on packets instead of individual features, and it compares each feature to features in similar regions of 
single-visit feature vectors instead of to the corresponding feature in other multi-visit algorithms.

@@

This algorithm seems to lead to good performance on multiclass SVMs, by picking out the few packet lengths/directions that 
distinguish target site visits. I used $\epsilon=100$ and $b=15$, and found these parameters to work well enough.
 In future, I would like to understand more rigorously why this algorithm works, what good choices of parameters are, how 
 this algorithm is similar or dissimilar to mutual information,
  and if there are
existing, well-studied alternatives that perform the same type of feature selection.

@@

In this project, I investigated the performance of several single- and multiclass SVMs on packet traces from visits to encrypted websites. 
I found that the recently proposed random-visit defense against traffic analysis does not hold up, since a SVM trained on 
dimensionality-reduced multi-visit data can achieve up to $70\%$ accuracy against this defense. In the future, I would like to better 
understand the feature extraction algorithm that I designed from a rigorous standpoint, and also study how to improve anomaly detection 
for the open-world setting of the encrypted traffic analysis problem.
